SNP Analysis

### 2019-12-24 Compare Models.Rmd

```{r, fig.width=8,fig.height=7}
kb.models = list.files(data)

@@

ggplot(kb.perf, aes(kb, mean, fill=kb)) + facet_grid(~alpha) + geom_col() + 
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
  geom_text(aes(label=n.coefs, y = 0.01)) + scale_fill_brewer(palette='Set1') + labs(x='',y='cv performance') +
  theme_minimal()

ggplot(kb.perf, aes(alpha, mean, fill=kb)) + facet_grid(~kb) + geom_col() + 
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
  scale_fill_brewer(palette='Set1') + labs(x='',y='cv performance') +
  geom_text(aes(label=n.coefs, y = 0.01)) + 
  theme_minimal()

@@

## LOO Training

@@

rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb)

ggplot(melt(rc9, id.vars='kb'), aes(variable, value, group=kb)) + 
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Training ROC (alpha=0.9)', x='', y='%') + theme_minimal()
```

@@

val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb)

ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + 
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation ROC (alpha=0.9)', x='', y='%') + theme_minimal()

@@

```{r fig.height=7, fig.width=8}
aucs = bind_rows(rc9 %>% mutate(cohort = 'Training'), val.auc %>% mutate(cohort = 'Validation'))

ggplot(aucs, aes(kb, AUC, group=cohort)) + 
  geom_bar(aes(fill=cohort), stat='identity', position='dodge') + 
  geom_text(aes(label=round(AUC,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='AUC training & validation', x='', y='AUC') + theme_minimal()
```

# Training LOO Performance

@@

## Training LOO alpha=0.9

@@

do.call(grid.arrange, c(plist, top='Training LOO alpha=0.9'))
```

@@

  pg.samp %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% 
    left_join(pg.samp %>% group_by(Status) %>% tally, by='Status') %>%
    dplyr::rename('Total' = 'n') %>%
    mutate_at(vars(Low,Moderate,High), funs(./Total)) 
})
names(per.samp.train) = list.files(data)

plist = purrr::map(names(per.samp.train), function(kb) {
  df = per.samp.train[[kb]]

  ggplot(melt(df, id.vars=c('Status','Total')), aes(variable, value, group=Status, fill=Status)) + 
    ylim(0,1) + geom_col(position='dodge') + 
    geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) + 
    labs(title=unique(kb), x='', y='ratio') + theme_minimal()
})
names(plist) = names(per.samp.train)

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Training predictions alpha=0.9'))

@@

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Validation'))

>>>>

```{r, fig.width=8,fig.height=7}
dir.create('plots/parameters')
kb.models = list.files(data)

@@

p = ggplot(kb.perf, aes(kb, mean, fill=kb)) + facet_grid(~alpha) + ylim(0,1) + 
  geom_col() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
  #geom_text(aes(label=n.coefs, y = 0.01)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(x='',y='classification accuracy', title = 'Cross-Validation Classification Accuracy', subtitle='Alpha values:') +
  theme_minimal() + theme(text = element_text(size=12), axis.text.x  = element_text(hjust = 0.5, angle=90))
p
ggsave(filename='plots/parameters/cv_accuracy.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)
```

@@

## LOO Discovery

@@

rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
  dplyr::select(-Threshold)

p = ggplot(melt(rc9, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) + 
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Discovery ROC', subtitle = 'alpha=0.9', x='', y='%') + 
  theme_minimal() + theme(text=element_text(size=12))
p
ggsave(filename='plots/parameters/training_kb_auc.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)

@@

val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
  dplyr::select(-Threshold)

ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) +
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation ROC (alpha=0.9)', x='', y='%') + theme_minimal()

@@

```{r fig.height=7, fig.width=8}
aucs = bind_rows(rc9 %>% mutate(Cohort = 'Discovery'), val.auc %>% mutate(Cohort = 'Validation'))

p = ggplot(aucs, aes(kb, AUC, group=Cohort)) + ylim(0,1) + 
  geom_bar(aes(fill=Cohort), stat='identity', position='dodge') + 
  geom_text(aes(label=round(AUC,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='AUC discovery & validation', x='', y='AUC') + 
  theme_minimal() + theme(text = element_text(size=12))
p
ggsave(filename='plots/parameters/tv_auc_kb_compare.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)

# Discovery LOO Performance

@@

## Discovery LOO alpha=0.9

@@

do.call(grid.arrange, c(plist[c('15kb','50kb','100kb','500kb')], top='Discovery LOO alpha=0.9'))
```

@@

  pg.samp %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% 
    left_join(pg.samp %>% group_by(Status) %>% tally, by='Status') %>%
    dplyr::rename('Total' = 'n') %>%
    mutate_at(vars(Low,Moderate,High), funs(./Total)) %>% mutate(kb = kb)
})
names(per.samp.train) = list.files(data)

plist = purrr::map(names(per.samp.train), function(kb) {
  df = melt(per.samp.train[[kb]], id.vars=c('Status','Total','kb')) %>% as_tibble %>% mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor'))
  ggplot(df, aes(Status, value, group=Status, fill=variable)) + 
    ylim(0,1) + geom_bar(stat = 'identity')  + 
    geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5)) + 
    scale_fill_manual(values =  BarrettsProgressionRisk:::riskColors(), name = 'Class') +
    labs(title=unique(kb), x='', y='sample ratio') + theme_minimal() + theme(text = element_text(size = 12), legend.position = 'none')
})
names(plist) = names(per.samp.train)

p = do.call(grid.arrange, c(plist[c('15kb','50kb','100kb','500kb')], top='Discovery Cohort Classifications (alpha=0.9)'))

ggsave(filename='plots/parameters/kb_classes.png', plot = p, width = 6, height = 12, units = 'in', dpi = 300)

# x = do.call(bind_rows, per.samp.train) 
# x = melt(x, id.vars=c('Status','Total', 'kb')) %>% as_tibble %>% mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor'), kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
#   ggplot(x, aes(Status, value, group=Status, fill=variable)) + facet_wrap(kb~Status, scale='free_x', nrow=2) + 
#     ylim(0,1) + geom_bar(stat = 'identity')  + 
#     geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5)) + 
#     scale_fill_manual(values =  BarrettsProgressionRisk:::riskColors(), name = 'Class') +
#     labs(title=unique(kb), x='', y='sample ratio') + theme_minimal() + theme(text = element_text(size = 12), legend.position = 'none')

@@

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title=tt, x='', y='ratio') + theme_minimal()
})
do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Validation'))

plist = purrr::map(names(val.counts), function(kb) {
  df = melt(val.counts[[kb]], id.vars=c('Status','kb')) %>% as_tibble %>% mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor'))
  ggplot(df, aes(Status, value, group=Status, fill=variable)) + 
    ylim(0,1) + geom_bar(stat = 'identity')  + 
    geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5)) + 
    scale_fill_manual(values =  BarrettsProgressionRisk:::riskColors(), name = 'Class') +
    labs(title=unique(kb), x='', y='sample ratio') + theme_minimal() + theme(text = element_text(size = 12), legend.position = 'none')
})
names(plist) = names(val.counts)

p = do.call(grid.arrange, c(plist[c('15kb','50kb','100kb','500kb')], top='Validation Cohort Classifications (alpha=0.9)'))

ggsave(filename='plots/parameters/kb_val_classes.png', plot = p, width = 6, height = 12, units = 'in', dpi = 300)

### 2019-12-24 Paper_resubmit.Rmd

val.info %>% group_by(Status) %>% dplyr::summarise(
  `Total Patients` = length(unique(Patient)),
  `Total Samples` = length(Samplename),
`max follow-up` = max(`Weeks Follow Up`/52),
`min follow-up` = min(`Weeks Follow Up`/52),
  `mean follow-up` = mean(`Weeks Follow Up`/52),
  `sd follow-up` = sd(`Weeks Follow Up`/52)
)

val.info %>% group_by(Status, Pathology) %>% tally

@@

val.info %>% filter(Samplename %in% failedQC$samplename)

@@

second %>% filter(Status == 'NP' & Rule == '1') %>% ungroup  %>%
  dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.)))

>>>>

val.info

@@

@@

second %>% filter(Status == 'NP' & Rule == '1') %>% ungroup  %>%
  dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.))) 
```

## Check COX

```{r}
cox.cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/cox_5e6_all/50kb/'
file = paste(cox.cache.dir, 'model_data.Rdata', sep='/')
load(file, verbose=T)

file = paste(cox.cache.dir, "all.pt.alpha.Rdata" , sep='/')
load(file, verbose=T)

do.call(grid.arrange, c(plots, nrow=2))

file = paste0(cox.cache.dir, '/loo_0.9.Rdata')
load(file, verbose=T)

rocCox = pROC::roc(Status ~ FittedRR, data=pg.samp, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocCox) + labs(title = 'Cox model ROC')

first = pg.samp %>% group_by(Patient) %>% arrange(Endoscopy.Year) %>% filter(Endoscopy.Year == Endoscopy.Year[1])
rocCox2 = pROC::roc(Status ~ FittedRR, data = first, auc = T, ci = T, of='thresholds', transpose=T)
roc.plot(rocCox2) + labs(title = 'Cox model ROC, first endo')

last = filter(pg.samp, Endoscopy.Year == Final.Endoscopy)
rocCox3 = pROC::roc(Status ~ FittedRR, data = last, auc = T, ci = T, of='thresholds', transpose=T)
roc.plot(rocCox3) + labs(title = 'Cox model ROC, last endo')

rocPath = pROC::roc(Status ~ Pathology, data=pg.samp, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocPath) + labs(title = 'Pathology status ROC')

cor.test(pg.samp$FittedRR, as.integer(pg.samp$Pathology))

### 2019-12-24 snp/SNP_Prediction_Analysis.Rmd

```{r, echo=F}
sm = sample.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% 
  group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) ) 
sm$Pathology = factor(sm$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T)

@@

ggplot(qcdata, aes(Status, ASCAT.SCA.ratio)) + ylim(0,0.05) + 
  facet_grid(~SampleType, scales='free_x') +
  geom_jitter(width=0.1, aes(color=Status)) + 
  geom_hline(data=dplyr::rename(anno, Status=SampleType), aes(yintercept = Ratio, linetype=Status), color='grey42') +
  geom_label(data=dplyr::rename(anno, Status=SampleType), aes(x='NP', y=Ratio, label=label), alpha=0.5)

@@

ggplot(filter(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, fill=Timepoint, group=Timepoint)) + geom_jitter(width=0.2) +
  geom_boxplot(alpha=0.5) + facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')

@@

p = ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio)) + facet_wrap(~SampleType, nrow=1) + 
  geom_point(aes(color=ASCAT.SCA.ratio < x | Purity > y), show.legend = F) + 
  geom_vline(xintercept=y, color='grey') + 
  geom_hline(yintercept=x, color='grey') +
  labs(title='Purity~SCA.Ratio') + theme(text = element_text(size=12))
p
ggsave('plots/snp/purity_v_sca.png', plot=p, height=4, width=4, units = 'in', dpi=300)

@@

ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
  geom_hline(yintercept = x, color='red') + labs(title='Barretts only')

@@

tmp = qcdata.endo %>% filter(!is.na(Ploidy_mean)) %>% mutate(Ploidy_mean = factor(round(Ploidy_mean)))
p = ggplot(tmp, aes(Ploidy_mean, ASCAT.SCA.ratio_mean, group=Ploidy_mean)) + 
  facet_grid(~SampleType) +
  geom_jitter(width=0.2, aes(color=Ploidy_mean)) + 
  geom_boxplot(outlier.colour = NA, alpha=0.5) + 
  geom_hline(yintercept = 0.01, color='grey39') + labs(title='Merged per endoscopy', x='Mean ploidy') + 
  theme(text=element_text(size=12), legend.position = 'none')
ggsave('plots/snp/sca_v_ploidy.png', plot=p, height=4, width=4, units = 'in', dpi=300)

@@

@@

Despite that, we still see a fairly consistent high/low risk prediction for NP vs P.  However, HGD in NP patients predicts at a rate consistent with progression.
```{r echo=F}
preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T))
rp = plot.risk.by.path(preds) 
rp

@@

### ROC 
```{r}
roc = pROC::roc(Status ~ Prob, data=bind_rows(preds, normals %>% mutate(Status = 'NP')), auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples')
pROC::ci.auc(roc$auc)
pROC::coords(roc, 'best', transpose=F)
ggsave(filename = 'plots/snp/all_samp_roc.png', plot = roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE samples only')
pROC::ci.auc(rocBE$auc)
pROC::coords(rocBE, 'best', transpose=F)
ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE samples only') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

@@

@@

rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocEX) + labs(title='BE samples', subtitle='Path status (HGD/IMC)')
pROC::ci.auc(rocEX)
ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE samples', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocpp) + labs(title='BE samples', subtitle='excluding low SCA samples')
ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE samples', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Exclude all of the above
rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocexall) + labs(title='BE samples', subtitle='excluding low SCA & WGD NP')
ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE samples', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Timepoint 1 only
roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct1) + labs(title='BE samples', subtitle='Timepoint T1')
ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE samples', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Timepoint 2 only
roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct2) + labs(title='BE samples', subtitle='Timepoint T2')
ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE samples', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

>>>>

```{r, echo=F}
sample.info = sample.info %>% mutate(Pathology = recode(Pathology, IMC = 'EAC'))

sm = sample.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% 
  group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) ) %>% 
  mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))

@@

p = ggplot(qcdata %>% filter(grepl('Normal',SampleType)), aes(Purity,ASCAT.SCA.ratio)) + 
  ylim(0,0.02) +
  geom_point(color='steelblue' ) + labs(title='Blood/Gastric Normal ASCAT values')
p
ggsave('plots/snp/normal_sca_purity.png', p, height=4, width=4, units='in')

ggplot(qcdata %>% filter(!grepl('Normal',SampleType)), aes(Purity,ASCAT.SCA.ratio)) + 
  #ylim(0,0.5) +
  geom_point(color='steelblue' ) + 
  geom_point(data = qcdata %>% filter(!grepl('Normal',SampleType) & ASCAT.SCA.ratio < 0.01 & Purity < 0.98), color='lightgreen') +
  labs(title='BE biopsy ASCAT values')

# 
# ggplot(qcdata, aes(Status, ASCAT.SCA.ratio)) + ylim(0,0.05) + 
#   facet_grid(~SampleType, scales='free_x') +
#   geom_jitter(width=0.1, aes(color=Status)) + 
#   geom_hline(data=dplyr::rename(anno, Status=SampleType), aes(yintercept = Ratio, linetype=Status), color='grey42') +
#   geom_label(data=dplyr::rename(anno, Status=SampleType), aes(x='NP', y=Ratio, label=label), alpha=0.5)

@@

# ggplot(filter(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, fill=Timepoint, group=Timepoint)) + 
#   geom_jitter(width=0.2) + geom_boxplot(alpha=0.5) + 
#   facet_wrap(~Status, scales='free_x', ncol=2) + 
#   labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')

@@

p = ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio)) + facet_wrap(~SampleType, nrow=1) + 
  geom_point(aes(color=ASCAT.SCA.ratio < x & Purity > y), show.legend = F) + 
  geom_vline(xintercept=y, color='grey') + 
  geom_hline(yintercept=x, color='grey') +
  labs(title='Purity~SCA.Ratio') + theme(text = element_text(size=12))
p
ggsave('plots/snp/purity_v_sca.png', plot=p, height=4, width=4, units = 'in', dpi=300)

@@

ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
  geom_hline(yintercept = x, color='grey39', linetype = 'dashed') + labs(title='Barretts biopsies')

@@

p = ggplot(qcdata.endo, aes(Purity_mean, ASCAT.SCA.ratio_mean)) + facet_grid(~SampleType) +
  geom_point(aes(color=ASCAT.SCA.ratio_mean < 0.01 & Purity_mean > 0.95 ), show.legend = F) + 
  labs(title='Merged biopsies per-timepoint', y='mean(ASCAT.SCA.ratio)', x='mean(ASCAT purity)')
p
ggsave('plots/snp/sca_purity_per_tp.png', p, width=4, height=4, units='in')

tmp = qcdata.endo %>% filter(!is.na(Ploidy_mean)) %>% mutate(Ploidy_mean = factor(round(Ploidy_mean)))
p = ggplot(tmp, aes(Ploidy_mean, ASCAT.SCA.ratio_mean, group=Ploidy_mean)) + 
  facet_grid(~SampleType) +
  geom_jitter(width=0.2, aes(color=Ploidy_mean)) + 
  geom_boxplot(outlier.colour = NA, alpha=0.5) + 
  geom_hline(yintercept = 0.01, color='grey39') + labs(title='Merged per endoscopy', x='Mean ploidy') + 
  theme(text=element_text(size=12), legend.position = 'none')
p
ggsave('plots/snp/sca_v_ploidy.png', plot=p, height=4, width=4, units = 'in', dpi=300)

@@

patient.info = patient.info %>% mutate(Pathology = recode(Pathology, 'IMC' = 'EAC'))

@@

Despite that, we still see a fairly consistent high/low risk prediction for NP vs P.  However, HGD in NP patients predicts at a rate consistent with progression.
```{r echo=F}
preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))
rp = plot.risk.by.path(preds) 
rp

@@

### ROC 
```{r}
# roc = pROC::roc(Status ~ Prob, data=bind_rows(preds, normals %>% mutate(Status = 'NP')), auc=T, ci=T, of='thresholds',transpose=T)
# roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples')
# pROC::ci.auc(roc$auc)
# pROC::coords(roc, 'best', transpose=F)
# ggsave(filename = 'plots/snp/all_samp_roc.png', plot = roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies')
pROC::ci.auc(rocBE$auc)
pROC::coords(rocBE, 'best', transpose=F)
ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

@@

pred.qc %>% filter(!lowsca | wgd.NP) %>% ungroup %>% dplyr::select(PatientID) %>% distinct

@@

rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)')
pROC::ci.auc(rocEX)
ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples')
ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Exclude all of the above
rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP')
ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Timepoint 1 only
roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1')
ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Timepoint 2 only
roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2')
ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
